\NAME: A framework for \\ Efficient Intrusion Detection\\ in the Internet of Things

### 2014-10-21 paper.tex
completed draft rewrite of background/design

  \item[Interactive] communication allows algorithms to exchange information to
  reach a consensus. In \cite{krontiris2009cooperative}, the authors first
  present a theoretical foundation to analyze cooperative algorithms. Given
  these foundations they continue to present an algorithm consisting of 5
  phases. It essentially implements a voting system based on the Guy Fawkes
  protocol\cite{anderson1998new} to identify an intruder. It enables the
  exchange of suspected intruder information and allows distributed
  authentication of those \emph{votes}. Based on these authenticated votes, a
  distributed decision can be made about the commonly identified intruder.

@@

The DSL enables researchers to formally describe their ID algorithms, without
focusing on any specific language or platform. Given such formal descriptions
of ID algorithms, a code generator is now able to combine these, while avoiding
the problems arising from the sequential patterns/problems presented in
listings \ref{alg:id-algo-application} and \ref{alg:id-algo-pattern}. It can
also target different languages and platforms, while reusing a software library
that provides e.g. message parsing, network access wrapping,\dots

\subsection{DSL}

The DSL component of \NAME has two major goals and one minor: (1) allow the
description of ID algorithms in a formal way and (2) avoid that those
descriptions contain \emph{non-fusible} logic and/or data. Besides these (3)
the usability of the language is also taken into account.

The first goal is covered by the construction of the language itself. Given a
description of an ID algorithm using the DSL, the algorithm can be
unambiguously interpreted. The DSL borrows a lot of its syntax from C-like
languages, dropping only a small subset of its capabilities, but replacing
those with other higher-level constructs taken from object-oriented (OO)
languages, aspect-orientation,\dots

The most important omission from the C syntax are basic loops. Loops are very
generic and make it hard to identify if they deal with the same logic and/or
data. Fusing these loops across different algorithms is often non-trivial. In
\NAME's DSL, other constructs have been introduced to lift the definition of
these loops to a more functional level: scheduling and scoping of the execution
of functions, reacting to events and pattern matching are three examples that
allow describing the actual meaning of loops in stead of simply the loop
itself. It sometimes requires rethinking the reason behind a technical loop,
but often results in a much clearer description.

To deal with common data, the domain provides the notion of \emph{nodes}. As
most, if not all, ID algorithms targeting the IoT require some notion of the
other nodes or devices in the vicinity, the DSL makes this the primary
information carrier. All functional constructs, mentioned above, typically
operate on the collection of known nodes, or a functionally defined subset of
that. Each algorithm also has the possibility to \emph{extend} nodes, adding
its own properties needed for its execution.

Based on a C-like syntax the language is also very accessible. To further
nurture the usability, syntactic sugar from OO languages is also adopted.
Because not everything should be implemented in such a high-level language, it
also provides a way to \emph{import} external functionality. In the first place
this will include a standard library of functions that can be called, but it
can include any custom built software component. Finally it tries to limit the
need for type information and performs as much type inference as possible. As
it aims to be language that describes algorithms on a functional level it seeks
to safeguard its users from these \emph{technicalities}. The fact that the
opportunities of such a formal description include generation of code in a
variety of languages, each with possibly different types, also results the need
for a more functional approach to types.

Listing \ref{lst:watchdog} illustrates some of the features mentioned above,
using a small but representative example of an algorithm: a
watchdog\cite{mishra2004intrusion}. Using a heartbeat, nodes can broadcast a
message at a regular interval, allowing other nodes to monitor its
availability. Missing heartbeats could indicate problems, such as physical
capturing of a certain node.
\lstdefinelanguage{foo-lang}{
  emph={module,const,from,import,extend,with,after,do,function,@every,having},
  emphstyle={\textbf},
  morecomment=[l]{//},
  moredelim=[is][]{@@}{@@},       % temp solution to allow keywords without emph
  moredelim=[is][\emph]{!!}{!!}   % temp solution to hightlight #atoms
}

@@

To convert the DSL based ID algorithms into production-ready code, we propose
the use of a code generator. Although many approaches here are applicable, like
template based text expansion, XSLT transformations or even using the
CodeDom\cite{dollard2004code}, we believe that for the level of functional
transformations that need to be performed, a richer model is required.

@@

- incoming message parsing
- outgoing message grouping
- execution of schedules
- standard library: high-level conceptual functions (e.g. now(), sha1,...)

>>>>

  \item[Interactive] communication allows algorithms to exchange information to
  reach a consensus. In \cite{krontiris2009cooperative}, the authors first
  present a theoretical foundation to analyze cooperative algorithms. Given
  these foundations they continue to present an algorithm consisting of 5
  phases. It essentially implements a voting system-based on the Guy Fawkes
  protocol\cite{anderson1998new} to identify an intruder. It enables the
  exchange of suspected intruder information and allows distributed
  authentication of those \emph{votes}. Based on these authenticated votes, a
  distributed decision can be made about the commonly identified intruder.

@@

The DSL enables researchers to formally describe their ID algorithms, without
focusing on any specific language or platform. Given such formal descriptions
of ID algorithms, a code generator is now able to combine these, while avoiding
the problems arising from the patterns/problems presented in listings
\ref{alg:id-algo-application} and \ref{alg:id-algo-pattern}. It can also target
different languages and platforms, while reusing a software library that
provides e.g. message parsing, network access wrapping,\dots

\subsection{The Meaning of Loops}

From the design patterns we learned that loops are a major contributor to the
problem of algorithm fusibility. Loops in program code are technical constructs
and in fact often hide the actual intended functionality.

To illustrate this, we take the example of parsing network messages. Each
algorithm deals with parsing in the same way: loop over the bytes in the
payload and look for a pattern that identifies what the algorithm is looking
for. This is in fact a technical implementation choice. The underlying
functional goal is to react when such pattern is present in a message. The loop
has \emph{activated} this functionality, but the real goal is to
\emph{passively} react on an event, not \emph{actively} look for it.

This is how \NAME removes the need for loops: by introducing event handling,
constructions with loops are transformed to their functional meaning, allowing
much easier identification and fusion of common functionalities.

The same goes for actively polling some variable until it reaches a certain
threshold. This also can be transformed to an event, eliminating the need for a
technical loop-construction.

\subsection{A DSL for Nodes}

The second problem that could be observed from the patterns, was that of common
data. All algorithms deal with information about neighboring devices. To
centralize this information \NAME provides a unified \texttt{nodes} concept as
part of its DSL and exposes it as an object to the implementor.

Nodes can be extended with properties only available to the algorithm that
defines them. This way only one instance for every known devices will be
present in memory. Finally, nodes can define a scope for some function to be
executed in. This way, nodes also offer a way to iterate all known devices,
again without the need for a loop.

\subsection{Unified Messaging}

Sending and receiving of network messages is tightly integrated with the
previously mentioned concepts of event handling and the nodes domain. An
underlying software framework will handle all interaction with the network,
triggering functions by means of events when new messages are parsed and
exposing an API to send messages through the nodes domain. This approach gives
the framework complete control over the way messages are sent, allowing
optimized marshaling, grouping,\dots

Listing \ref{lst:watchdog} illustrates some of the features mentioned above,
using a small but representative example of an algorithm: a
watchdog\cite{mishra2004intrusion}. Like a heartbeat, nodes can broadcast a
message at a regular interval, allowing other nodes to monitor its
availability. Missing heartbeats could indicate problems, such as physical
capturing of a device.

\lstdefinelanguage{foo-lang}{
  emph={module,const,from,import,extend,with,after,do,function,@every,having},
  emphstyle={\textbf},
  morecomment=[l]{//},
  moredelim=[is][]{@@}{@@},       % temp solution to allow keywords without emph
  moredelim=[is][\emph]{!!}{!!}   % temp solution to hightlight #atoms
}

@@

To convert the DSL-based ID algorithms into production-ready code, we propose
the use of a code generator. Although many approaches here are applicable, like
template-based text expansion, XSLT transformations or even using the
CodeDom\cite{dollard2004code}, we believe that for the level of functional
transformations that need to be performed, a richer model is required.

@@

We already touched upon the underlying software library when discussing the
unified messaging framework: incoming messages are parsed, outgoing messages
grouped. Besides these examples, other common functionality also isn't
generated. The execution of functions based on a schedule or the implementation
of higher level conceptual functions, such as \texttt{now} and \texttt{sha1},
are calls into the standard software library that supports the code generator.
